### 2019-06-17 report/chapter03.Rnw
Results section update

% LaTeX file for Chapter 03
<<'preamble03',include=FALSE>>=
library(knitr)
opts_chunk$set(
    fig.path='figure/ch03_fig', 
    self.contained=FALSE,
    cache=FALSE
) 
@

<<echo=FALSE>>=
library(knitr)
opts_chunk$set(
    fig.path='figure/ch02_fig',
    self.contained=FALSE,
    cache=TRUE
)
@

<<echo=FALSE>>=
library(knitr)
opts_chunk$set(fig.path='figure/ch02_fig',
               echo=TRUE, message=FALSE,
               fig.width=8.1, fig.height=3,
               out.width='\\textwidth-3cm',
               message=FALSE, fig.align='center',
               background="gray98", tidy=FALSE, #tidy.opts=list(width.cutoff=60),
               cache=TRUE
)
options(width=74)
@

<<echo=FALSE>>= 
rm(list = ls())
PATH_HOME = path.expand("~") # user home
PATH = file.path(PATH_HOME, 'Data/PubBias')
PATH2 = file.path(PATH_HOME, 'PubBias')
FILE = 'cochrane_2018-06-09.csv'
PATH_DATA = file.path(PATH, 'data')
PATH_CODE = file.path(PATH2, 'code')
PATH_RESULTS = file.path(PATH2, 'results')
PATH_FIGURES = file.path(PATH_RESULTS, 'figures')

@@

file.bin <- "pb.bin.RData"
if (file.exists(file.path(PATH_RESULTS, file.bin))) {
  load(file.path(PATH_RESULTS, file.bin))
} else {
  meta.bin <- meta.bin.complete(data.ext, min.study.number = 10, sig.level = 0.05, sm = "OR")
  save(meta.bin, file =  file.path(PATH_RESULTS, file.bin))
}

file.cont <- "pb.cont.RData"
if (file.exists(file.path(PATH_RESULTS, file.cont))) {
  load(file.path(PATH_RESULTS, file.cont))
} else {
  meta.cont <- meta.cont.complete(data.ext, min.study.number = 10, sig.level = 0.05)
  save(meta.cont, file =  file.path(PATH_RESULTS, file.cont))
}

file.meta <- "meta.RData"
if (file.exists(file.path(PATH_RESULTS, file.meta))) {
  load(file.path(PATH_RESULTS, file.meta))
} else {
  meta <- pb.meta.merge(meta.bin, meta.cont)
  save(meta, file =  file.path(PATH_RESULTS, file.meta))
}

load(file.path(PATH_RESULTS, file = "data.processed.RData"))

require(biostatUZH)
require(tidyverse)
require(meta)
require(metasens)
require(gridExtra)
require(xtable)
@

@@

% 
% \section{Meta-analysis}
% The data at hand is composed of over 400,000 results with balanced sample size $n > 24$. The significance of the p-values of treatment effect estimates was calculated for all results in the dataset with acceptable sample size and outcome measures as odds ratios, peto odds ratios, risk ratios, risk differences and mean and standardized mean differences. 
% 
% %Significance over time? Chunck 1
% 
% The results that can be pooled are analysed by fixed-effects meta-analysis, random-effects meta-analysis and random-effects meta-analysis with Hartung-Knapp method.
% 
% \vspace{0mm}
% In the following, the results are combined with the results of their corresponding meta-analysis if possible (i.e. if studies were not unique in the dataset with respect to their experimental setup). The significance of the treatment effect (primary significance) in a single result is then compared to the significance of the pooled treatment effect (secondary significance). 
% In Figure \ref{primary.secondary.significance}, the overall fraction of studies with significant and non-significant meta-analysis results is shown. The meta-analysis method used to pool the treatment effect estimates and assessing significance is indicated on the right hand side of the Figure. The colors indicate the overall fraction of rejected or accepted null-hypotheses of no treatment effect.
% 
% \begin{figure}
% <<echo=FALSE, warning=FALSE>>=
% #plot(sig.meta.c)
% @
% \caption{Overall fraction of studies whose treatment effect estimate was significant when pooled by means of meta-analysis. The fractions have been calculated
% by fixed-effects, random-effects and Hartung and Knapp adjusted random-effects meta-analysis}
% \label{primary.secondary.significance}
% \end{figure}
% 
% What has been found is that even the meta-analysis that for all meta-analysis methods, the amount of secondary significance is higher than the amount of primary significance. Thus more primarily non-significant results can be attached to a secondarily significant result than vice-versa.
% 
% \vspace{0mm}
% To illustrate this, the fraction of studies with significant primary treatment effect and primary non-significant treatment effect can be plotted separately (Figure \ref{primary.secondary.significance.sep.sig}). 
% 
% \begin{figure}
% <<echo=FALSE, warning=FALSE>>=
% #grid.arrange(sig.meta.plot1, nonsig.meta.plot, ncol= 1)
% @
% \caption{Overall fraction of studies whose treatment effect estimate was significant when pooled by means of meta-analysis, separated by significance of study treatment effect estimate. The fractions have been calculated
% by fixed-effects, random-effects and Hartung and Knapp adjusted random-effects meta-analysis}
% \label{primary.secondary.significance}
% \end{figure}
% 
% It can be seen that the extent of overlap between primary and secondary significance varies between meta-analysis. A considerable amount of both primary significant and primary non-significant results are overruled by meta-analysis hypothesis tests. 
% 
% 
% The amount of significant heterogeneity in the results used by the meta-analyses is \Sexpr{length(which(mly$sig.Q == 0))} and corresponds to \Sexpr{(length(which(mly$sig.Q == 0))/ dim(mly)[1])*100}\% of the total number of meta-analyses.
% 
% %Comparison of significance separated for significant Q? Chunck 2
% 
% %Proportion of significant pooled estimates based on proportion of single pooled estimates? Chunck 5

\section{Small study effects}

<<echo=FALSE>>=
p.samplesize.effect <- data %>% filter(total1 + total2 > 10 & total1 + total2 < 200) %>% 
  mutate(sample.size = total1 + total2, scaled.effect = abs(scale(effect, center = T, scale = T))) %>%
  group_by(sample.size) %>% 
  summarize(median.effect = median(scaled.effect, na.rm = T)) %>% 
  ggplot(aes(x = sample.size, y = median.effect)) + geom_point() + theme_bw() + xlab("sample size") + ylab("median absolute normalized effect size")

#levels(data$outcome.measure.new) <- c("Mean Difference", "Std. Mean Difference", "Odds Ratio", "Risk Ratio")

p.samplesize.effect.sep <- data %>% filter(total1 + total2 > 10 & total1 + total2 < 100) %>% 
  filter(outcome.measure.new == "Risk Ratio" | outcome.measure.new == "Odds Ratio" | outcome.measure.new == "Mean Difference" |
           outcome.measure.new == "Std. Mean Difference") %>% 
  group_by(outcome.measure.new) %>% 
  mutate(sample.size = total1 + total2, scaled.effect = abs(scale(effect, center = T, scale = T))) %>%
  group_by(sample.size, outcome.measure.new) %>% 
  summarize(median.effect = median(scaled.effect, na.rm = T)) %>% 
  ggplot(aes(x = sample.size, y = median.effect, group = outcome.measure.new)) + geom_point() + facet_wrap(~outcome.measure.new, scales = "free") + 
  theme_bw() + xlab("sample size") + ylab("median absolute normalized effect size")
@

To provide an overview over the abundance of small study effects in the dataset, first it is shown how median absolute effect size decreases with increasing sample size of the results (Figure \ref{effect.samplesize}). 
%In some sense, this is the same idea as for a funnel plot, by depicting the size of effects relative to their variance.

\vspace{0mm}
A clear trend of decrease of absolute effect size with increasing sample size (i.e. smaller variance) is visible. 
%It is particularly substantive from very small trials ($n$ = 10) to medium sample size ($n$ = 100) and afterwards it evens off. With increasing sample size, there are fewer results, therefore, the variation between medians increases. 
All effects are normalized by subtracting the mean effect size of the dataset and dividing through the standard deviation. Note that various types of outcome measures are included, such as mean difference and risk ratios, and are normalized with respect to all effects.

\begin{figure}
<<echo=FALSE>>=
plot(p.samplesize.effect)
@
\caption{Median of the absolute value of the normalized effect size plotted against the total sample size.}
\label{effect.samplesize}
\end{figure}

The median absolute normalized effect size can be visualized for the different outcome measures separately (Figure \ref{effect.samplesize.separated}). The plot confirms the trend of median effect size decrease towards lower sample size (e.g. for Risk Ratios: decrease $> 0.1$ standard deviations from 10 to 50 study participants). Instead of normalizing the effects 
%(i.e. subtracting mean and dividing through the standard deviation) 
for all effect sizes, the effects are normalized with respect to the effects of the same outcome measures.

\begin{figure}
<<echo=FALSE>>=
plot(p.samplesize.effect.sep)
@
\caption{Median of the absolute value of the normalized effect size plotted against the total sample size, separated for outcome measures.}
\label{effect.samplesize.separated}
\end{figure}

@@

\subsection{Small Study Effect Tests}

<<echo=FALSE, warning=FALSE>>=
#Test Results: Binary
test.bin <- meta.bin %>% ungroup() %>% summarize(egger.test = mean(egger.test),
                                                          schwarzer.test = mean(schwarzer.test),
                                                          rucker.test = mean(rucker.test),
                                                          harbord.test = mean(harbord.test),
                                                          peter.test = mean(peter.test))
test.bin <- test.bin %>% gather(key = "test.type", value = "mean")

p.bin <- meta.bin %>% ungroup() %>% 
  select(egger.test, schwarzer.test, rucker.test, harbord.test, peter.test) %>% 
  gather(key = "test.type", value = "null.hypothesis") %>%  
  mutate(null.hypothesis = factor(ifelse(null.hypothesis == 1, "rejected", "not rejected"))) %>% 
  ggplot(aes(x = test.type, fill = null.hypothesis)) + geom_bar() + coord_flip() + 
  theme_bw() + xlab(label = NULL) +
  annotate("text", x = test.bin$test.type, y = 1000, 
           label = paste(round(test.bin$mean, 2)*100, "% rejected"), 
           color = "white")

test.sig.bin <- meta.bin %>% filter(sig.fixef.bin == 1) %>% ungroup() %>% summarize(egger.test = mean(egger.test),
                                                              schwarzer.test = mean(schwarzer.test),
                                                              rucker.test = mean(rucker.test),
                                                              harbord.test = mean(harbord.test),
                                                              peter.test = mean(peter.test))
test.sig.bin <- test.sig.bin %>% gather(key = "test.type", value = "mean")

p1 <- meta.bin %>% ungroup() %>% filter(sig.fixef.bin == 1) %>% 
select(egger.test, schwarzer.test, rucker.test, harbord.test, peter.test) %>% 
  gather(key = "test.type", value = "null.hypothesis") %>% 
  mutate(null.hypothesis = factor(ifelse(null.hypothesis == 1, "rejected", "not rejected"))) %>% 
  ggplot(aes(x = test.type, fill = null.hypothesis)) + geom_bar() + coord_flip() + 
  theme_bw() + ggtitle("Significant Pooled Effects")+ xlab(label = NULL) + theme(legend.position="none") +
  annotate("text", x = test.sig.bin$test.type, y = 1750, 
           label = paste(round(test.sig.bin$mean, 2)*100, "% rejected"), 
           color = "white")

test.nonsig.bin <- meta.bin %>% filter(sig.fixef.bin == 0) %>% ungroup() %>% summarize(egger.test = mean(egger.test),
                                                                                                schwarzer.test = mean(schwarzer.test),
                                                                                                rucker.test = mean(rucker.test),
                                                                                                harbord.test = mean(harbord.test),
                                                                                                peter.test = mean(peter.test))
test.nonsig.bin <- test.nonsig.bin %>% gather(key = "test.type", value = "mean")

p2 <- meta.bin %>% 
  filter(sig.fixef.bin == 0) %>% ungroup() %>% 
  select(egger.test, schwarzer.test, rucker.test, harbord.test, peter.test) %>% 
  gather(key = "test.type", value = "null.hypothesis") %>% 
  mutate(null.hypothesis = factor(ifelse(null.hypothesis == 1, "rejected", "not rejected"))) %>% 
  ggplot(aes(x = test.type, fill = null.hypothesis)) + geom_bar() + coord_flip() + 
  theme_bw() + ggtitle("Non-significant Pooled Effects")+ xlab(label = NULL) +theme(legend.position="none") +
  annotate("text", x = test.nonsig.bin$test.type, y = 650, 
           label = paste(round(test.nonsig.bin$mean, 2)*100, "% rejected"), 
           color = "white")

range.pb.difference.bin <- range(test.sig.bin$mean - test.nonsig.bin$mean)

#Test Results: Continuous
test.cont <- meta.cont %>% ungroup() %>% summarize(egger.test = mean(egger.test),
                                                            begg.test = mean(begg.test),
                                                            thomson.test = mean(thomson.test))

@@

p.cont <- meta.cont %>% ungroup() %>% 
  select(egger.test, thomson.test, begg.test) %>% 
  gather(key = "test.type", value = "null.hypothesis") %>% 
  mutate(null.hypothesis = factor(ifelse(null.hypothesis == 1, "rejected", "not rejected"))) %>% 
  ggplot(aes(x = test.type, fill = null.hypothesis)) + geom_bar() + coord_flip() + 
  theme_bw() + xlab(label = NULL) +
  annotate("text", x = test.cont$test.type, y = 750, 
           label = paste(round(test.cont$mean, 2)*100, "% rejected"), 
           color = "white")

test.sig.cont <- meta.cont %>% filter(sig.fixef.cont == 1) %>% ungroup() %>% summarize(egger.test = mean(egger.test),
                                                                                                begg.test = mean(begg.test),
                                                                                                thomson.test = mean(thomson.test))

test.sig.cont <- test.sig.cont %>% gather(key = "test.type", value = "mean")

p3 <- meta.cont %>% 
  filter(sig.fixef.cont == 1) %>% ungroup() %>% 
  select(egger.test, thomson.test, begg.test) %>% 
  gather(key = "test.type", value = "null.hypothesis") %>% 
  mutate(null.hypothesis = factor(ifelse(null.hypothesis == 1, "rejected", "not rejected"))) %>% 
  ggplot(aes(x = test.type, fill = null.hypothesis)) + geom_bar() + coord_flip() + 
  theme_bw() + ggtitle("Significant Pooled Effects")+ xlab(label = NULL) + theme(legend.position="none") +
  annotate("text", x = test.sig.cont$test.type, y = 600, 
           label = paste(round(test.sig.cont$mean, 2)*100, "% rejected"), 
           color = "white")

test.nonsig.cont <- meta.cont %>% filter(sig.fixef.cont == 0) %>% ungroup() %>% summarize(egger.test = mean(egger.test),
                                                                                                   begg.test = mean(begg.test),
                                                                                                   thomson.test = mean(thomson.test))

test.nonsig.cont <- test.nonsig.cont %>% gather(key = "test.type", value = "mean")

p4 <- meta.cont %>% 
  filter(sig.fixef.cont == 0) %>% ungroup() %>% 
  select(egger.test, thomson.test, begg.test) %>% 
  gather(key = "test.type", value = "null.hypothesis") %>% 
  mutate(null.hypothesis = factor(ifelse(null.hypothesis == 1, "rejected", "not rejected"))) %>% 
  ggplot(aes(x = test.type, fill = null.hypothesis)) + geom_bar() + coord_flip() + 
  theme_bw() + ggtitle("Non-significant Pooled Effects")+ xlab(label = NULL) + theme(legend.position="none") +
  annotate("text", x = test.nonsig.cont$test.type, y = 100, 
           label = paste(round(test.nonsig.cont$mean, 2)*100, "% rejected"), 
           color = "white")

range.pb.difference.cont <- range(test.sig.cont$mean - test.nonsig.cont$mean)

########################################################################################################################################
########################################################################################################################################

#Agreement proportions of publication bias tests:

#Binary:
n.bin <- dim(meta.bin)[1]
meta.bin <- meta.bin %>%ungroup() %>%  mutate(egger.schwarzer = ifelse(egger.test == schwarzer.test, "agree", "disagree"),
                                        egger.peter = ifelse(egger.test == peter.test, "agree", "disagree"),
                                        egger.rucker = ifelse(egger.test == rucker.test, "agree", "disagree"),
                                        egger.harbord = ifelse(egger.test == harbord.test, "agree", "disagree"),
                                        schwarzer.peter = ifelse(schwarzer.test == peter.test, "agree", "disagree"),
                                        schwarzer.rucker = ifelse(schwarzer.test == rucker.test, "agree", "disagree"),
                                        schwarzer.harbord = ifelse(schwarzer.test == harbord.test, "agree", "disagree"),
                                        rucker.peter = ifelse(rucker.test == peter.test, "agree", "disagree"),
                                        rucker.harbord = ifelse(rucker.test == harbord.test, "agree", "disagree"),
                                        harbord.peter = ifelse(harbord.test == peter.test, "agree", "disagree"))

agreement.bin <- meta.bin %>% ungroup() %>% summarise(egger.schwarzer = sum(egger.schwarzer == "agree")/n(),
                                            egger.peter = sum(egger.peter == "agree")/n(),
                                            egger.rucker = sum(egger.rucker == "agree")/n(),
                                            egger.harbord = sum(egger.harbord == "agree")/n(),
                                            schwarzer.peter = sum(schwarzer.peter == "agree")/n(),
                                            schwarzer.rucker = sum(schwarzer.rucker == "agree")/n(),
                                            schwarzer.harbord = sum(schwarzer.harbord == "agree")/n(),
                                            rucker.peter = sum(rucker.peter == "agree")/n(),
                                            harbord.peter = sum(harbord.peter == "agree")/n())

correlation.bin <- meta.bin %>%ungroup() %>%  summarise(egger.schwarzer = cor(pval.egger.bin, pval.schwarzer.bin),
                                              egger.peter = cor(pval.egger.bin, pval.peter.bin),
                                              egger.rucker = cor(pval.egger.bin, pval.rucker.bin),
                                              egger.harbord = cor(pval.egger.bin, pval.harbord.bin),
                                              schwarzer.peter = cor(pval.schwarzer.bin, pval.peter.bin),
                                              schwarzer.rucker = cor(pval.schwarzer.bin, pval.rucker.bin),
                                              schwarzer.harbord = cor(pval.schwarzer.bin, pval.harbord.bin),
                                              rucker.peter = cor(pval.rucker.bin, pval.peter.bin),
                                              harbord.peter = cor(pval.harbord.bin, pval.peter.bin))

binary.tests.agreement <- rbind(agreement.bin, correlation.bin)
rownames(binary.tests.agreement) <- c("Test Agreement","P-value Correlation")

#Continuous:
n.cont <- dim(meta.cont)[1]
meta.cont <- meta.cont %>% ungroup() %>% mutate(thomson.egger = ifelse(thomson.test == egger.test, "agree", "disagree"),
                                          thomson.begg = ifelse(thomson.test == begg.test, "agree", "disagree"),
                                          egger.begg = ifelse(egger.test == begg.test, "agree", "disagree"))

agreement.cont <- meta.cont %>% ungroup() %>%  summarise(thomson.egger = sum(thomson.egger == "agree")/n(),
                                              thomson.begg = sum(thomson.begg == "agree")/n(),
                                              egger.begg = sum(egger.begg == "agree")/n())

correlation.cont <- meta.cont %>% ungroup() %>% summarise(thomson.egger = cor(pval.thomson.cont, pval.egger.cont),
                                                thomson.begg = cor(pval.thomson.cont, pval.begg.cont),
                                                egger.begg = cor(pval.egger.cont, pval.begg.cont))

>>>>

% LaTeX file for Chapter 03

%<<echo=FALSE>>=
% library(knitr)
% opts_chunk$set(
%      fig.path='figure/ch03_fig',
%      self.contained=FALSE,
%      cache=TRUE
%  )
%@

<<'preamble03',include=FALSE, echo=FALSE>>=
library(knitr)
opts_chunk$set(fig.path='figure/ch03_fig',
               echo=TRUE, message=FALSE,
               fig.width=8, fig.height=3,
               self.contained = FALSE,
               out.width='\\textwidth-3cm',
               message=FALSE, fig.align='center',
               background="gray98", tidy=FALSE, #tidy.opts=list(width.cutoff=60),
               cache=TRUE
)
options(width=74)
@

<<'Data03', echo=FALSE, warning=FALSE>>= 
rm(list = ls())
PATH_HOME = path.expand("~") # user home
PATH = file.path(PATH_HOME, 'Data/PubBias')
PATH2 = file.path(PATH_HOME, 'PubBias')
FILE = 'cochrane_2018-06-09.csv'
PATH_DATA = file.path(PATH, 'data')
PATH_CODE = file.path(PATH2, 'code')
PATH_RESULTS = file.path(PATH2, 'results')
PATH_FIGURES = file.path(PATH_RESULTS, 'figures')

@@

load(file.path(PATH_RESULTS, "meta_analyses_summary_complete.RData"))
load(file.path(PATH_RESULTS, "meta.bin.RData"))
load(file.path(PATH_RESULTS, "meta.cont.RData"))
load(file.path(PATH_RESULTS, "meta.surv.RData"))
load(file.path(PATH_RESULTS, "data_used_for_analysis.RData"))

data.ext2 <- pb.process2(data)
@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Small study effects}
<<echo=FALSE, warning=FALSE>>=
p.samplesize.z  <- data.ext2 %>% filter(total1 + total2 > 15 & total1 + total2 < 300) %>% ungroup() %>% 
  mutate(std.z = abs(z), sample.size = total1 + total2) %>%
  group_by(sample.size) %>% 
  summarize(median.effect = median(std.z, na.rm = T)) %>% 
  ggplot(aes(x = sample.size, y = median.effect)) + geom_point(size = 0.8) + 
  theme_bw() + xlab("sample size") + ylab("median absolute z-score") +
  xlim(c(15, 300)) + ylim(c(0, 0.25)) +
  scale_x_continuous(limits = c(15, 300), breaks = c(15, 100, 200, 300))

p.samplesize.effect.sep <- data %>% filter(total1 + total2 > 15 & total1 + total2 < 300) %>% 
  filter(outcome.measure.new == "Risk Ratio" | 
           outcome.measure.new == "Odds Ratio" | 
           outcome.measure.new == "Mean Difference" |
           outcome.measure.new == "Std. Mean Difference") %>% 
  group_by(outcome.measure.new) %>% 
  mutate(sample.size = total1 + total2, scaled.effect = abs(scale(effect, center = T, scale = T))) %>%
  group_by(sample.size, outcome.measure.new) %>% 
  summarize(median.effect = median(scaled.effect, na.rm = T)) %>% 
  ggplot(aes(x = sample.size, y = median.effect, group = outcome.measure.new)) + geom_point(size = .5) + facet_wrap(~outcome.measure.new, scales = "free") + 
  theme_bw() + xlab("sample size") + ylab("median absolute normalized effect size") +
  xlim(c(15, 300)) +
  scale_x_continuous(limits = c(15, 300), breaks = c(15, 100, 200, 300))
@

The median $z$-score for a given sample size of a trial is shown in Figure \ref{z.samplesize}. It is clearly visible that the absolute value decreases with increasing sample size, i.e. that the effect size is becoming smaller. The trend flattens off after $\sim$ sample size = 400 (not shown). 
-- Appendix --

\begin{figure}
<<echo=FALSE, warning=FALSE>>=
plot(p.samplesize.z)
@
\caption{Median of the absolute $z$-score plotted against the total sample size.}
\label{z.samplesize}
\end{figure}

Only for illustration, the same trend is reproduced in Figure \ref{effect.samplesize.separated} with a similar method, using the original effect size measures ``Odds Ratio'', ``Risk Ratio'', ``Mean Difference'' and ``Std. Mean Difference'' (the most commonly used in the dataset). The ``normalized effect'' is the original effect size, normalized with respect to all other effect sizes of the same measure (i.e. subtraction of mean and division through standard error of the mean).

\begin{figure}
<<echo=FALSE>>=
plot(p.samplesize.effect.sep)
@
\caption{Median of the absolute value of the normalized, original effect size plotted against the total sample size.}
\label{effect.samplesize.separated}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Small Study Effect Tests}
<<echo=FALSE, warning=FALSE>>=
sig.level <- 0.1
meta.f <- meta.f %>% 
	mutate(egger.test = ifelse(pval.egger < sig.level, 1, 0),
				 thompson.test = ifelse(pval.thompson < sig.level, 1, 0),
				 begg.test = ifelse(pval.begg < sig.level, 1, 0),

				 schwarzer.test = ifelse(pval.schwarzer < sig.level, 1, 0),
				 rucker.test = ifelse(pval.rucker < sig.level, 1, 0),
				 harbord.test = ifelse(pval.harbord < sig.level, 1, 0),
				 peter.test = ifelse(pval.peter < sig.level, 1, 0))

meta.bin <- meta.f %>% filter(outcome.type == "bin")
meta.cont <- meta.f %>% filter(outcome.type == "cont")
meta.surv <- meta.f %>% filter(outcome.type == "surv")

test.bin <- meta.bin %>% ungroup() %>% summarize(
																								 schwarzer.test = mean(schwarzer.test),
																								 rucker.test = mean(rucker.test),
																								 harbord.test = mean(harbord.test),
																								 peter.test = mean(peter.test))
test.bin <- test.bin %>% gather(key = "test.type", value = "mean")

p.bin <- meta.bin %>% ungroup() %>% 
	select(schwarzer.test, rucker.test, harbord.test, peter.test) %>% 
	gather(key = "test.type", value = "null.hypothesis") %>%  
	mutate(null.hypothesis = factor(ifelse(null.hypothesis == 1, "significant", "not significant"))) %>% 
	ggplot(aes(x = test.type, fill = null.hypothesis)) + geom_bar() + coord_flip() + 
	theme_bw() + xlab(label = NULL) + ggtitle("Binary Outcomes") + theme(legend.position = "top") +
	guides(fill=guide_legend(title=NULL))+
	annotate("text", x = test.bin$test.type, y = 500, 
					 label = paste(round(test.bin$mean, 2)*100, "% rejected"), 
					 color = "white")
#--------------------------------------------------------------------------------------------------------------------#

dat_text <- data.frame(
	label = paste(c(sum(meta.bin$harbord.test), sum(meta.bin$peter.test), 
									sum(meta.bin$rucker.test), sum(meta.bin$schwarzer.test)), "< 0.1,",
								round(c(mean(meta.bin$harbord.test), mean(meta.bin$peter.test), 
												mean(meta.bin$rucker.test), mean(meta.bin$schwarzer.test)),2)*100, "%"),
	test.type   = c("pval.harbord", "pval.peter", "pval.rucker", "pval.schwarzer"))

labels <- c(pval.harbord = "Harbord", pval.peter = "Peter", pval.rucker = "Rucker", pval.schwarzer = "Schwarzer")

p.dist.bin <- meta.bin %>% ungroup() %>% 
	select(pval.harbord, pval.peter, pval.rucker, pval.schwarzer) %>% 
	gather(key = "test.type", value = "p.value") %>% 
	ggplot(aes(x = p.value)) + geom_histogram(boundary = 0, binwidth = 0.1) + theme_bw() + facet_wrap(~test.type, labeller = labeller(test.type = labels)) + 
	geom_text(data = dat_text, mapping = aes(x = 0.5, y = 300, label = label), color = "black") + 
	theme(strip.text.x = element_text(size=10))
#--------------------------------------------------------------------------------------------------------------------#

#Survival Outcomes:
test.cont <- meta.cont %>% ungroup() %>% summarize(egger.test = mean(egger.test),
																									 begg.test = mean(begg.test),
																									 thompson.test = mean(thompson.test))

@@

p.cont <- meta.cont %>% ungroup() %>% 
	select(egger.test, thompson.test, begg.test) %>% 
	gather(key = "test.type", value = "null.hypothesis") %>% 
	mutate(null.hypothesis = factor(ifelse(null.hypothesis == 1, "significant", "not significant"))) %>% 
	ggplot(aes(x = test.type, fill = null.hypothesis)) + geom_bar() + coord_flip() + 
	theme_bw() + xlab(label = NULL) + ggtitle("Continuous Outcomes") + theme(legend.position = "top") +
	guides(fill=guide_legend(title=NULL))+
	annotate("text", x = test.cont$test.type, y = 150, 
					 label = paste(round(test.cont$mean, 2)*100, "% rejected"), 
					 color = "white")
#--------------------------------------------------------------------------------------------------------------------#

dat_text <- data.frame(
	label = paste(c(sum(meta.cont$begg.test), sum(meta.cont$egger.test), sum(meta.cont$thompson.test)), "< 0.1,",
								round(c(mean(meta.cont$begg.test), mean(meta.cont$egger.test), mean(meta.cont$thompson.test)),2)*100, "%"),
	test.type   = c("pval.begg", "pval.egger", "pval.thompson"))

labels <- c(pval.begg = "Begg Mazumdar", pval.egger = "Egger", pval.thompson = "Thompson Sharp")
p.dist.cont <- meta.cont %>% ungroup() %>% 
	select(pval.egger, pval.thompson, pval.begg) %>% 
	gather(key = "test.type", value = "p.value") %>% 
	ggplot(aes(x = p.value)) + geom_histogram(boundary = 0, binwidth = 0.1) + theme_bw() + 
	facet_wrap(~test.type, labeller = labeller(test.type = labels)) + 
	geom_text(data = dat_text, mapping = aes(x = 0.5, y = 90, label = label),  color = "black") + 
	theme(strip.text.x = element_text(size=10))
#--------------------------------------------------------------------------------------------------------------------#

#Survival Outcomes:
test.surv <- meta.surv %>% ungroup() %>% summarize(egger.test = mean(egger.test),
																									 begg.test = mean(begg.test),
																									 thompson.test = mean(thompson.test))

test.surv <- test.surv %>% gather(key = "test.type", value = "mean")

p.surv <- meta.surv %>% ungroup() %>% 
	select(egger.test, thompson.test, begg.test) %>% 
	gather(key = "test.type", value = "null.hypothesis") %>% 
	mutate(null.hypothesis = factor(ifelse(null.hypothesis == 1, "significant", "not significant"))) %>% 
	ggplot(aes(x = test.type, fill = null.hypothesis)) + geom_bar() + coord_flip() + 
	theme_bw() + xlab(label = NULL) + ggtitle("Continuous Outcomes") + theme(legend.position = "top") +
	guides(fill=guide_legend(title=NULL))+
	annotate("text", x = test.surv$test.type, y = 10, 
					 label = paste(round(test.surv$mean, 2)*100, "% rejected"), 
					 color = "white")
#--------------------------------------------------------------------------------------------------------------------#

dat_text <- data.frame(
	label = paste(c(sum(meta.surv$begg.test), sum(meta.surv$egger.test), sum(meta.surv$thompson.test)), 
	              "< 0.1,",
								round(c(mean(meta.surv$begg.test), mean(meta.surv$egger.test), mean(meta.surv$thompson.test)),2)*100, 
								"%"),
	test.type   = c("pval.begg", "pval.egger", "pval.thompson"))

labels <- c(pval.begg = "Begg Mazumdar", pval.egger = "Egger", pval.thompson = "Thompson Sharp")
p.dist.surv <- meta.surv %>% ungroup() %>% 
	select(pval.egger, pval.thompson, pval.begg) %>% 
	gather(key = "test.type", value = "p.value") %>% 
	ggplot(aes(x = p.value)) + geom_histogram(boundary = 0, binwidth = 0.1) + theme_bw() + 
	facet_wrap(~test.type, labeller = labeller(test.type = labels)) + 
	geom_text(data = dat_text, mapping = aes(x = 0.5, y = 11, label = label),  color = "black") + 
	theme(strip.text.x = element_text(size=10))
#--------------------------------------------------------------------------------------------------------------------#

#Test agreement
agree.bin <- meta.bin %>% mutate(n.sig = peter.test + rucker.test + harbord.test + schwarzer.test) %>% 
	group_by(n.sig) %>% count %>% filter(n.sig > 0) %>% 
	ggplot(aes(y = nn, x = n.sig)) + theme_bw() + geom_col() + 
  xlab("Number of significant tests") + ylab("count") + ggtitle("Binary Outcomes")
#--------------------------------------------------------------------------------------------------------------------#

agree.cont <- meta.cont %>% mutate(n.sig = egger.test + thompson.test + begg.test) %>% 
	group_by(n.sig) %>% count %>% filter(n.sig > 0) %>% 
	ggplot(aes(y = nn, x = n.sig)) + theme_bw() + geom_col() + 
  xlab("Number of significant tests") + ylab("count") + ggtitle("Continuous Outcomes")
#--------------------------------------------------------------------------------------------------------------------#

agree.surv <- meta.surv %>% mutate(n.sig = egger.test + thompson.test + begg.test) %>% 
	group_by(n.sig) %>% count %>% filter(n.sig > 0) %>% 
	ggplot(aes(y = nn, x = n.sig)) + theme_bw() + geom_col() + 
  xlab("Number of significant tests") + ylab("count") + ggtitle("Continuous Outcomes")
#--------------------------------------------------------------------------------------------------------------------#
@

The results of the test for small study effects for the meta-analyses fullfilling the criteria from chapter \ref{chapter01}, section \ref{sec:Processing}, are presented in Figures \ref{fig:test.bin}, \ref{fig:test.cont} and \ref{fig:test.surv}. The histogram of the $p$-values appears most often skewed to the right, indicating evidence for small study effects. Because different tests are applied depending on outcome type, the results are displayed separately for binary, continuous and survival outcome types (as previously defined in the dataset variable ``outcome.type'', see chapter \ref{ch:chapter01}, section \ref{sec:Processing}. \\
For continuous and survival outcomes, the names refer to:

\begin{itemize}
\item Egger's test, the weigthed linear regression test as described in section \ref{sec:Egger}
\item Thompson and Sharp's test, the weighted linear regression test adjusted for between-study heterogeneity, section \ref{sec:Thompson}
\item Begg and Mazumdar's test, the rank test described in section \ref{sec:Begg}
\end{itemize}

For binary outcomes, the names refer to:
\begin{itemize}
\item Harbord's test, the likelihood score based test (section \ref{sec:Harbord})
\item Peter's test, the weighted linear regression with inverse sample size as explanatory variable (study size proxy) described in section \ref{sec:Peter}
\item R\"ucker's test, the test based on the arcsine transformation of proportions, in combination with Thompson and Sharp's regression test (section \ref{sec:Rucker})
\item Schwarzer's test, the rank based test using the expected event counts computed with the hypergeometric distribution (section \ref{sec:Schwarzer})
\end{itemize}
